---
title: "Homework 3"
author: "Shaolei Ma"
date: "`r Sys.Date()`"
output: github_document
---

```{r libraries and initial settings, echo = F, message = F}
library(tidyverse)
library(ggridges)
library(patchwork)

library(p8105.datasets) # for loading datasets

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6, # width / height
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

```{r}
data("instacart") # import dataset
```

The data set `instacart` contains `r nrow(instacart)` observations and `r ncol(instacart)` variables related to the information of `r instacart |> pull(order_id) |> unique() |> length()` orders of `r instacart |> pull(user_id) |> unique() |> length()` users and they all belong to the train set. Key information includes the time of the order (`order_dow`, `order_hour_of_day`) and the location of the product (`aisle_id`, `aisle`, `department_id`, `department`).

 - How many aisles are there, and which aisles are the most items ordered from?
 
 There are `r instacart |> pull(aisle_id) |> unique() |> length()` aisles. The number of items ordered from each aisle is illustrated below:
 
 `r instacart |> count(aisle) |> arrange(desc(n)) |> head() |> knitr::kable()`
 
 So the top six aisles from which the most items ordered are: `r instacart |> count(aisle) |> arrange(desc(n)) |> head() |> pull(aisle)`.

 - Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.
```{r}
instacart |> 
  count(aisle) |> 
  arrange(desc(n)) |>
  filter(n > 20000) |> 
  ggplot(aes(x = aisle, y = n)) +
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 45)) # rotate the x-axis labels
```
 
 - Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.
```{r}
instacart |> 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |> 
  count(aisle, product_name, sort = T) |> 
  group_by(aisle) |> 
  top_n(3, n) |> # top 3 products within each group
  knitr::kable()
```
 
 - Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).
```{r}
instacart |> 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |> 
  group_by(product_name,order_dow) |> 
  summarise(mean_hour = mean(order_hour_of_day)) |> 
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  ) |> 
  knitr::kable()
```
 
 